
\section{Protocol Architecture}

The Horizon Protocol is organized into layered components with clearly defined responsibilities and upgrade boundaries. This separation isolates external dependencies from core accounting logic and limits the surface area affected by upgrades.

\subsection{Layer Model}

\begin{center}
\begin{tabular}{lll}
\toprule
\textbf{Layer} & \textbf{Components} & \textbf{Upgrade Scope} \\
\midrule
External & Underlying yield tokens, Pragma Oracle & Outside protocol control \\
Core & SY, PT, YT, Market & Upgradeable via governed proxy \\
Periphery & Router, Factory, MarketFactory & Upgradeable via governed proxy \\
\bottomrule
\end{tabular}
\end{center}

\section{Token Flow}

The protocol implements a reversible asset flow centered around a deterministic split of standardized yield into principal and yield claims.

\subsection{Forward Flow (Deposit and Split)}

\begin{enumerate}
    \item A user deposits an underlying yield-bearing token into the SY contract.
    \item The SY contract mints a standardized yield token representing the deposited asset.
    \item The user deposits SY into the YT contract, which atomically mints paired PT and YT tokens such that $1\,\text{SY} = 1\,\text{PT} + 1\,\text{YT}$.
\end{enumerate}

\subsection{Reverse Flow (Redemption)}

\begin{enumerate}
    \item Before expiry, a user may redeem exactly one PT and one YT to recover one SY, reversing the split.
    \item At expiry, PT redeems deterministically 1:1 for SY, while YT expires with zero value.
    \item The user withdraws the underlying asset from the SY contract, receiving principal plus any accrued yield.
\end{enumerate}

\section{Token System}

The protocol defines four token types representing distinct economic claims. All tokens implement the ERC-20 interface pattern and impose no transfer restrictions.

\begin{center}
\begin{tabularx}{\textwidth}{l >{\raggedright\arraybackslash}X l l}
\toprule
\textbf{Token} & \textbf{Economic Role} & \textbf{Minting Mechanism} & \textbf{Transferability} \\
\midrule
SY & Standardized yield wrapper & Permissionless via deposit & Freely transferable \\
PT & Principal claim (redeemable at expiry) & Minted via YT contract & Freely transferable \\
YT & Yield claim (pre-expiry) & Minted via YT contract & Freely transferable \\
LP & Liquidity provider share & Minted by Market contract & Freely transferable \\
\bottomrule
\end{tabularx}
\end{center}

\section{Core Protocol Invariants}

The following invariants are safety-critical and must hold across all valid state transitions of the protocol.

\begin{enumerate}
    \item \textbf{Principal Protection}: Each PT redeems deterministically for exactly one unit of SY at maturity.
    \item \textbf{Conservation}: Prior to expiry, the combined value of PT and YT corresponds to one unit of SY.
    \item \textbf{Price Convergence}: The AMM enforces monotonic convergence of PT price toward redemption value as expiry approaches.
    \item \textbf{Yield Isolation}: All yield generated by the underlying asset accrues exclusively to YT holders prior to expiry.
\end{enumerate}
