
\section{Conclusion}

Horizon Protocol introduces a yield tokenization framework on Starknet, enabling users to separate principal and yield exposure for yield-bearing assets.

\paragraph{Key contributions.}
\begin{itemize}
\item \textbf{Starknet-native yield tokenization.} A PT/YT model adapted to Starknet's execution and arithmetic model, providing the first yield tokenization infrastructure on the network.

\item \textbf{Cairo-native numerical design.} High-precision fixed-point arithmetic via the cairo-fp 64.64 library, enabling accurate transcendental function computation for logit-based pricing.

\item \textbf{Native oracle integration.} Pragma TWAP-based exchange rate tracking for yield-bearing assets, with configurable staleness thresholds and watermark protection.

\item \textbf{Logit-based AMM implementation.} A complete implementation of the Pendle-style curve with deterministic price convergence at expiry.
\end{itemize}


\section{Core Guarantees}

The following guarantees define the protocol's economic correctness and distinguish it from lending-based fixed-rate systems.

\begin{center}
\begin{tabular}{l l l}
\toprule
\textbf{Guarantee} & \textbf{Mechanism} & \textbf{Enforcement} \\
\midrule
PT redemption at par & PT $\to$ 1 SY after expiry & Contract logic \\
Token conservation & PT + YT = SY always holds & Arbitrage + contract \\
Deterministic convergence & $r_{scalar} \to \infty$ as $\tau \to 0$ & AMM mathematics \\
Supply equality & PT.supply = YT.supply & Contract assertion \\
Index monotonicity & $I_{global}$ only increases & Watermark logic \\
\bottomrule
\end{tabular}
\end{center}

Together, these guarantees ensure that PT holders receive exactly 1 SY at expiry regardless of market conditions, YT holders receive all accrued yield, and LPs face predictable convergence dynamics.


\section{Design Trade-offs}

The following trade-offs were intentionally chosen to balance correctness, complexity, and deployability for a v1 protocol.

\begin{center}
\begin{tabularx}{\textwidth}{l l >{\raggedright\arraybackslash}X}
\toprule
\textbf{Trade-off} & \textbf{Choice} & \textbf{Rationale} \\
\midrule
Upgradeability vs immutability & Upgradeable & Enables bug fixes; requires admin trust \\
Curve complexity vs simplicity & Full logit curve & Capital efficiency justifies complexity \\
Gas cost vs precision & cairo-fp 64.64 & Financial math requires high precision \\
Single-block vs multi-block execution & Single-transaction & Reduces state complexity; oracle TWAP mitigates manipulation \\
\bottomrule
\end{tabularx}
\end{center}

\paragraph{Note on execution model.}
All protocol operations complete within a single transaction. Multi-block execution patterns (e.g., commit-reveal) were not implemented. Oracle manipulation resistance relies on Pragma's TWAP mechanism rather than protocol-level multi-block designs.


\section{Out-of-Scope Enhancements}

The following enhancements are out of scope for this specification and are not commitments. They are listed to clarify possible future directions.

\begin{center}
\begin{tabular}{l l}
\toprule
\textbf{Enhancement} & \textbf{Purpose} \\
\midrule
Timelock for upgrades & Allow users to exit before upgrade execution \\
Multisig requirement & Reduce single-key compromise risk \\
Additional SY adapters & Support more yield-bearing assets \\
Cross-market routing & Optimize multi-hop trade execution \\
Governance token & Enable decentralized protocol control \\
\bottomrule
\end{tabular}
\end{center}

These items may be addressed in future protocol versions but are explicitly excluded from this specification.


\section{Specification Status}

This document specifies Horizon Protocol v1.0.

\begin{center}
\begin{tabular}{l l l}
\toprule
\textbf{Component} & \textbf{Status} & \textbf{Coverage} \\
\midrule
Token model (SY, PT, YT) & Complete & Mint, redeem, transfer, yield accounting \\
AMM mechanism & Complete & Logit curve, fee decay, liquidity operations \\
Redemption flows & Complete & Pre-expiry and post-expiry paths \\
Oracle integration & Complete & Pragma TWAP, staleness, watermark \\
Governance model & Complete & Upgradeability, RBAC, emergency procedures \\
Risk analysis & Complete & Threat model, attack surfaces, mitigations \\
Parameter specification & Complete & Bounds, defaults, configuration \\
\bottomrule
\end{tabular}
\end{center}

\vspace{1em}

\noindent\textbf{This document constitutes the complete technical specification for Horizon Protocol v1.0. Any behavior not described herein is considered undefined.}

