
\section*{Glossary}
\label{sec:glossary}

This glossary defines key terms used throughout this specification. Terms are ordered alphabetically.

\begin{center}
\begin{tabularx}{\textwidth}{l >{\raggedright\arraybackslash}X}
\toprule
\textbf{Term} & \textbf{Definition} \\
\midrule

\textbf{AMM} &
Automated Market Maker; a smart contract that provides liquidity and determines prices algorithmically based on pool reserves and a pricing function. \\

\textbf{cairo-fp} &
Cairo fixed-point arithmetic library using 64.64 format (64 integer bits, 64 fractional bits), enabling high-precision transcendental function computation. \\

\textbf{Expiry ($T$)} &
Unix timestamp (seconds) at which PT becomes redeemable for exactly 1 SY and YT stops accruing yield. \\

\textbf{Implied Rate ($r$)} &
The annualized yield rate inferred from the current PT/SY exchange rate, assuming PT converges to 1 SY at expiry. Computed as $r = -\ln(P_{PT}) / \tau_y$. \\

\textbf{Logit} &
Transformation $\ln(p/(1-p))$ mapping pool proportions in $(0,1)$ to the real line $(-\infty, +\infty)$, used to derive the AMM exchange rate. \\

\textbf{LP Token} &
Token representing proportional ownership of a PT/SY liquidity pool; entitles holder to pro-rata share of reserves and accumulated fees. \\

\textbf{Market} &
Smart contract implementing the PT/SY AMM, responsible for swaps, liquidity provision, and price discovery. \\

\textbf{Minimum Liquidity} &
Small amount of LP tokens (typically $10^3$) permanently locked on first deposit to prevent share inflation attacks. \\

\textbf{PT (Principal Token)} &
Token representing a claim on principal that redeems for exactly 1 unit of SY at or after expiry. Trades at a discount before expiry reflecting the implied yield. \\

\textbf{PY} &
A matched pair of PT and YT tokens (1 PY = 1 PT + 1 YT), which together are economically equivalent to 1 SY before expiry. \\

\textbf{PY Index ($I_{global}$)} &
Monotonically non-decreasing cumulative index tracking yield accrual for YT holders. Updated from oracle on state-changing operations. \\

\textbf{Rate Anchor ($r_{anchor}$)} &
Dynamic offset parameter that ensures implied yield continuity across successive trades; updated after each swap to preserve rate consistency. \\

\textbf{Rate Scalar ($r_{scalar}$)} &
Time-dependent scaling parameter that controls AMM price sensitivity to changes in pool composition. Increases toward infinity as expiry approaches, forcing $P_{PT} \to 1$. \\

\textbf{RBAC} &
Role-Based Access Control system governing privileged contract operations (upgrades, pausing, parameter changes). \\

\textbf{SY (Standardized Yield)} &
Wrapper token that standardizes heterogeneous yield-bearing assets into a uniform accounting and redemption interface. Implements ERC-20 with deposit/redeem functions. \\

\textbf{Time to Expiry ($\tau$)} &
Seconds remaining until expiry: $\tau = T - t_{now}$. Often normalized to years: $\tau_y = \tau / Y$ where $Y = 31{,}536{,}000$. \\

\textbf{TWAP} &
Time-Weighted Average Price; oracle mechanism that averages prices over a time window to mitigate short-term manipulation. \\

\textbf{Underlying Asset} &
The original yield-bearing asset (e.g., nstSTRK, wstETH) that is wrapped by SY. Distinct from SY itself. \\

\textbf{WAD} &
Fixed-point scaling factor equal to $10^{18}$, used for all external-facing token amounts, rates, and prices. \\

\textbf{Watermark} &
Accounting mechanism that prevents the yield index from decreasing, protecting YT holders' accrued yield from negative oracle movements. \\

\textbf{YT (Yield Token)} &
Token representing the exclusive right to all yield generated by the underlying asset from mint time until expiry. Yield is claimable via \texttt{claim\_interest}. \\

\bottomrule
\end{tabularx}
\end{center}


\section*{Symbol Reference}

\begin{center}
\begin{tabular}{l l l}
\toprule
\textbf{Symbol} & \textbf{Name} & \textbf{Units} \\
\midrule
$P_{PT}$ & PT price in SY terms & Dimensionless (0 to 1) \\
$P_{YT}$ & YT price in SY terms & Dimensionless (0 to 1) \\
$r$ & Implied annual rate & Per-year (dimensionless) \\
$\tau$ & Time to expiry & Seconds \\
$\tau_y$ & Time to expiry (annualized) & Years \\
$T$ & Expiry timestamp & Unix seconds \\
$Y$ & Seconds per year & 31,536,000 \\
$I_{global}$ & Global yield index & WAD-scaled ratio \\
$I_{user}$ & User yield index snapshot & WAD-scaled ratio \\
$r_{scalar}$ & Rate scalar parameter & WAD-scaled \\
$r_{anchor}$ & Rate anchor parameter & WAD-scaled \\
$p$ & PT proportion in pool & Dimensionless (0 to 1) \\
\bottomrule
\end{tabular}
\end{center}

\vspace{2em}

\begin{center}
\rule{0.5\textwidth}{0.4pt}
\end{center}

\vspace{1em}

\begin{center}
\textit{End of Specification}
\end{center}

\vspace{1em}

\noindent\textbf{Document Version:} v2.0\\
\textbf{Last Updated:} December 2025\\
\textbf{License:} BUSL-1.1 (converts to GPL-3.0 on 2028-12-19)

\vspace{1em}

\noindent\textit{This document is a technical specification and does not constitute investment, legal, or financial advice. Users should conduct independent analysis before interacting with the protocol.}
